**Reproducible Research Course - Project 1**
===============================================

## Section 1 - Loading and Preprocessing the Data
Data is to be loaded into R as a data.frame.  
At this point, no further processing is required.

```{r}
df <- read.csv("activity.csv")
```

## Section 2 - What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset:  
1. Make a histogram of the total number of steps taken each day  
Sum the number of steps by each day and generate a histogram.

```{r}
sumbydate <- aggregate.data.frame(df$steps, by=list(df$date), FUN=sum)
library(ggplot2)
qplot(sumbydate$x, xlab = "Steps per day", ylim = c(0,10), main = 
              "Total number of steps taken each day")
```

2. Calculate and report the mean and median total number of steps taken per day

```{r}
mean <- mean(sumbydate$x, na.rm = TRUE)
median <- median(sumbydate$x, na.rm = TRUE)
```

The mean number of steps taken per day is `r mean` and the median is `r median`.

## Section 3 - What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)  

```{r}
time_series <- tapply(df$steps, df$interval, mean, na.rm = TRUE)
plot(row.names(time_series), time_series, type = "l", xlab = "Interval", 
     ylab = "Average across all Days", main = "Average steps taken")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max_interval <- which.max(time_series)
names(max_interval)
```


## Section 4 - Imputing missing values

1. Calculate and report the total number of missing values in the dataset 
(i.e. the total number of rows with NAs)

```{r}
sum(is.na(df))
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
The strategy does not need to be sophisticated. For example, you could use 
the mean/median for that day, or the mean for that 5-minute interval, etc.
- I will replace the NAs with the mean for that 5- minute interval.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
rmNA_df <- df

for (i in which(sapply(rmNA_df, is.numeric))) {
     for (j in which(is.na(rmNA_df[, i]))) {
         rmNA_df[j, i] <- mean(rmNA_df[rmNA_df[, "interval"] == rmNA_df[j, "interval"], i],  na.rm = TRUE)
    }
}

str(df)
str(rmNA_df)
```

4. Make a histogram of the total number of steps taken each day and Calculate 
and report the mean and median total number of steps taken per day. 
Do these values differ from the estimates from the first part of the assignment?
What is the impact of imputing missing data on the estimates of the total
daily number of steps?

```{r}
sumbydate <- aggregate.data.frame(rmNA_df$steps, by=list(rmNA_df$date), FUN=sum)
library(ggplot2)
qplot(sumbydate$x, xlab = "Steps per day", ylim = c(0,10), main = 
              "Total number of steps taken each day")

mean <- mean(sumbydate$x, na.rm = TRUE)
median <- median(sumbydate$x, na.rm = TRUE)
```

The mean number of steps taken per day is `r mean` and the median is `r median`.
Adjusting the NA values did not change the mean, but made the median equal to the mean.

## Section 5 - Are there differences in activity between weekdays and weekends?

For this part the weekdays function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.


2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
j <- as.character(rmNA_df$date)
k <- as.Date.character(j)
weekday <- weekdays(k)
weekday <- as.data.frame(weekday)
New_df <- cbind(rmNA_df,weekday)
New_df$weekday <- as.character(New_df$weekday)

for (i in 1:nrow(New_df)) {                                         
    if (New_df[i,]$weekday %in% c("Saturday","Sunday")) {             
        New_df[i,]$weekday<-"weekend"                                 
    }
    else{
        New_df[i,]$weekday<-"weekday"                                 
    }
}
stepsByDay <- aggregate(steps ~ interval + weekday, data = New_df, FUN = "mean")


names(stepsByDay) <- c("interval", "weekday", "steps")
library(lattice)
xyplot(steps ~ interval | weekday, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```
